---
title: "MTH 490 Project"
author: "Joseph Wuest"
date: "2023-11-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Import Necessary Packages
```{r}
library("readxl")
library("tidyverse")
library("ggplot2")
library("ggiraph")
library("lubridate")
```

# Import statistical d ata, clean it so that only goalies with 250 or more shots
# against are visible. Also, convert any necessary columns from class character 
# to class numeric. This involves removing commas.

```{r}
hockey_data <- read_excel("Summary.xlsx")

hockey_data$SA <- as.numeric(gsub(",", "", hockey_data$SA))
hockey_data$Svs <- as.numeric(gsub(",", "", hockey_data$Svs))
hockey_data$GA <- as.numeric(gsub(",", "", hockey_data$GA))
hockey_data$`Sv%` <- as.numeric(hockey_data$`Sv%`)



cleaned_data <- hockey_data %>% select(1, 5:8, 11:14, 16)
cleaned_data <- cleaned_data %>% filter(SA > 250)
cleaned_data
```


# Use full data set to gather average statistics, such as a league average save
# percentage.

```{r}
total_shots <- sum(hockey_data$SA)
total_saves <- sum(hockey_data$Svs)
total_goals <- sum(hockey_data$GA)

avg_sv_pct <- total_saves/total_shots
```


# Compute Sharpe Ratio, add it to the data frame


```{r}
sv_pct_sd <- sd(na.omit(hockey_data$'Sv%'))
cleaned_data$SR <- (cleaned_data$`Sv%` - avg_sv_pct)/sv_pct_sd
```


# Merge salary data to the cleaned data, must rename data headers first


```{r}
salary_data <- read_excel("Salary Data NHL.xlsx")
colnames(salary_data)[1] <- "Player"
colnames(salary_data)[2] <- "Cap Hit"
colnames(salary_data)[3] <- "Salary"
```

```{r}
complete_data <- merge(cleaned_data, salary_data, by="Player")
```


# Plot salary against sharpe ratio


```{r}
sr_to_salary <- ggplot(complete_data, )+
                geom_point(aes(x = SR, y = Salary/1000000), color = "black", size = 1)+
                geom_vline(xintercept = 0, color = "red", linetype  = "33")+
                xlab("Modified Sharpe Ratio")+
                ylab("Salary (in Millions of dollars)")
                
sr_to_salary
```

```{r}
complete_data$"TOI (min)" <- gsub('.{3}$', '', complete_data$TOI)
complete_data$`TOI (min)` <- as.numeric(complete_data$`TOI (min)`)

toi_to_salary <- ggplot(complete_data, )+
                geom_point(aes(x = `TOI (min)`, y = Salary/1000000))+
                xlab("Time on the Ice")+
                ylab("Salary (in Millions of dollars)")
                
toi_to_salary
```

```{r}

sv_pct_to_salary <- ggplot(complete_data, )+
                geom_point(aes(x = `Sv%`, y = Salary/1000000))+
                xlab("Save Percentage")+
                ylab("Salary (in Millions)")
                
sv_pct_to_salary
```


```{r}
sv_pct_to_toi <- ggplot(complete_data, )+
              geom_point(aes(x = SR , y = Salary/`TOI (min)`*`Sv%`))
sv_pct_to_toi
```


```{r}
complete_data$`W%` <- complete_data$W/(complete_data$W+complete_data$L)
sr_to_w_pct <- ggplot(complete_data, aes(x = SR, y = `W%`))+
            geom_point() + 
            geom_smooth(method='lm', formula = y ~ x, se = FALSE, color = "green", linetype = "2212")+
            xlab("Modified Sharpe Ratio")+
            ylab("Win Percentage")
sr_to_w_pct
```
```{r}
w_pct_to_salary <- ggplot(complete_data, )+
                geom_point(aes(x = `W%`, y = Salary))
w_pct_to_salary
```


```{r}
high_salary_data <- complete_data %>%
                    filter(Salary > 3000000)
high_salary_to_sr <- ggplot(high_salary_data, aes(x = SR, y = Salary/1000000))+
                  geom_point()+
                  xlab("Sharpe Ratio")+
                  ylab("Salary (in Millions of dollars)")+
                  geom_smooth(method='lm', formula = y ~ x, se = FALSE, color = "darkblue")+
                  geom_vline(xintercept = 0, linetype = "dashed", color = "red")+
                  xlim(-.5, .5)
                  
high_salary_to_sr
```





